The {TestIdeals} package for \emph{Macaulay2}

### 2019-04-18 Fsing/TestIdealsPaper/TestIdeals.tex
Changes 1 and 3 asked by referee:M2 and summands
Following referee's comments on the paper, I cited M2 in the bibliography and I modified the issue between terms and summands, writing it as referee suggests.
Changes to doc of testIdeal

\begin{document}
\title{The  {TestIdeals} package for \emph{Macaulay2}}
\author[Alberto F.\ Boix et al.]{Alberto F.\ Boix}
\address{Department of Mathematics, Ben-Gurion University of the Negev, Beer-Sheva 8410501, Israel}
\email{fernanal@post.bgu.ac.il}
\thanks{A.F.~Boix was supported by Israel Science Foundation (grant No. 844/14) and Spanish Ministerio de Econom\'ia y Competitividad MTM2016-7881-P}

@@

\begin{definition}[\cite{HernandezTeixeiraWittFrobeniusPowers}]
Let  $I\subseteq R$ be an ideal.
\begin{enumerate}
 \item[(a)] If $n$ is a positive integer with base $p$ expansion  $n=d_0 + d_1 p +  \dots + d_r p^r$, we define
\[ I^{[n]}=I^{d_0} (I^{d_1})^{[p]} \dots  (I^{d_r})^{[p^r]}.\]
 \item[(b)] If $t$ is a nonnegative rational number of the form $t = a/p^e$, we define  $I^{[t]} = {\big(I^{[a]}\big)}^{[1/p^e]}.$
 \item[(c)] If $t$ is any nonnegative rational number, and $\{a_n/p^{e_n}\}_{n\geq 1}$ is a sequence of rational numbers converging to $t$ from above, we define $I^{[t]}$
 to be the stable value of the non-decreasing chain of ideals $\{I^{[a_n/p^{e_n}]}\}_{n\geq 1}$.
\end{enumerate}
\end{definition}

@@

\begin{example}
Let $R=\ZZ/p\ZZ[a,b,c,d]$, and consider the ideals $\mathfrak{m}=(a,b,c,d)$ and
$$I= (a,b) \cap (a,c) \cap (c,d) \cap (c+d, a^3+b d^2).$$
Then $R/I$ is a $3$-dimensional non-Cohen-Macaulay ring.
Matlis Duality applied to $H^2_{\mathfrak{m}} (R/I)$ with its natural Frobenius map
yields a $p^{-1}$-linear map $U$ on $\Ext^2_R(R/ I, R)$.
%$\Ext^2(R/ I, R) \xrightarrow{U} \Ext^2(R/I, R)$.

@@

The performance of the algorithm can be improved if the ring of interest is
nice enough. If the ring is Cohen-Macaulay, then setting \texttt{AssumeCM =>
true} (the default value is \texttt{false}) lets the algorithm check the Frobenius action only on top cohomology
(which is typically much faster, as explained above).
%The default value for the option \texttt{AssumeCM} is \texttt{false}.
%Of
%course, telling the algorithm to assume the ring is Cohen-Macaulay when it
%is not can lead to an incorrect answer if the non-injective Frobenius
%occurs in a lower degree. For an example of this see the documentation.
When studying a reduced ring,  setting \texttt{AssumeReduced => true} (the default value) avoids
computing the bottom local cohomology, and when studying a normal ring, setting
\texttt{AssumeNormal => true} (the default is \texttt{false}) avoids computing the bottom two local
cohomologies.

@@

\begin{definition}
Suppose that $R$ is a Cohen-Macaulay ring and that $T_{R} :
\omega_{R^{1/p}} \rightarrow \omega_{R}$ is the canonical dual to the
Frobenius. We say that $R$ has \emph{$F$-rational singularities}, of simply that \emph{$R$ is $F$-rational}, if there
are no non-zero proper submodules $M$ of $\omega_{R}$ such that
$T_{R}(M^{1/p}) \subseteq M$.
\end{definition}

@@

The algorithm used by \texttt{isFRational} first checks if the ring is Cohen-Macaulay, unless the option \texttt{AssumeCM} is set to \texttt{true}. If the ring
is not Cohen-Macaulay then \texttt{false} is returned. Next, the algorithm
computes the test module $M \subseteq \omega_{R}$ and checks to see if
$\omega_{R} \subseteq M$;  see the next section for the description of a test module.

@@

In \cite{KatzmanParameterTestIdealOfCMRings} it is shown that as $R/I$ is Cohen-Macaulay, $u$
can be taken as the generator of the cyclic module $(I^{[p]}:I) \cap (\Omega^{[p]}:\Omega)/I^{[p]}$
where $\Omega \subseteq R$ is an ideal whose image in $R/I$ is a canonical module of that ring.

>>>>

\begin{document}
\title{The  {TestIdeals} package for \emph{Macaulay2}}
\author[Alberto F.\ Boix et al.]{Alberto F.\ Boix}
\address{Department of Mathematics, Ben-Gurion University of the Negev, Beer-Sheva 8410501, Israel}
\email{fernanal@post.bgu.ac.il}
\thanks{A.F.\,Boix was supported by Israel Science Foundation (grant No. 844/14) and Spanish Ministerio de Econom\'ia y Competitividad MTM2016-7881-P}

@@

\begin{definition}[\cite{HernandezTeixeiraWittFrobeniusPowers}]
Let  $I\subseteq R$ be an ideal.
\begin{enumerate}
 \item[(a)] If $n$ is a positive integer with base $p$ expansion  $n=d_0 + d_1 p +  \dots + d_r p^r$, we define
\[ I^{[n]}=I^{d_0} (I^{d_1})^{[p]} \cdots  (I^{d_r})^{[p^r]}.\]
 \item[(b)] If $t$ is a nonnegative rational number of the form $t = a/p^e$, we define  $I^{[t]} = {\big(I^{[a]}\big)}^{[1/p^e]}.$
 \item[(c)] If $t$ is any nonnegative rational number, and $\{a_n/p^{e_n}\}_{n\geq 1}$ is a sequence of rational numbers converging to $t$ from above, we define $I^{[t]}$
 to be the stable value of the non-decreasing chain of ideals $\{I^{[a_n/p^{e_n}]}\}_{n\geq 1}$.
\end{enumerate}
\end{definition}

@@

\begin{example}
Let $R=\ZZ/p\ZZ[a,b,c,d]$, and consider the ideals $\mathfrak{m}=(a,b,c,d)$ and
$$I= (a,b) \cap (a,c) \cap (c,d) \cap (c+d, a^3+b d^2).$$
Then $R/I$ is a $3$-dimensional non-Cohen--Macaulay ring.
Matlis Duality applied to $H^2_{\mathfrak{m}} (R/I)$ with its natural Frobenius map
yields a $p^{-1}$-linear map $U$ on $\Ext^2_R(R/ I, R)$.
%$\Ext^2(R/ I, R) \xrightarrow{U} \Ext^2(R/I, R)$.

@@

The performance of the algorithm can be improved if the ring of interest is
nice enough. If the ring is Cohen--Macaulay, then setting \texttt{AssumeCM =>
true} (the default value is \texttt{false}) lets the algorithm check the Frobenius action only on top cohomology
(which is typically much faster, as explained above).
%The default value for the option \texttt{AssumeCM} is \texttt{false}.
%Of
%course, telling the algorithm to assume the ring is Cohen--Macaulay when it
%is not can lead to an incorrect answer if the non-injective Frobenius
%occurs in a lower degree. For an example of this see the documentation.
When studying a reduced ring,  setting \texttt{AssumeReduced => true} (the default value) avoids
computing the bottom local cohomology, and when studying a normal ring, setting
\texttt{AssumeNormal => true} (the default is \texttt{false}) avoids computing the bottom two local
cohomologies.

@@

\begin{definition}
Suppose that $R$ is a Cohen--Macaulay ring and that $T_{R} :
\omega_{R^{1/p}} \rightarrow \omega_{R}$ is the canonical dual to the
Frobenius. We say that $R$ has \emph{$F$-rational singularities}, of simply that \emph{$R$ is $F$-rational}, if there
are no non-zero proper submodules $M$ of $\omega_{R}$ such that
$T_{R}(M^{1/p}) \subseteq M$.
\end{definition}

@@

The algorithm used by \texttt{isFRational} first checks if the ring is Cohen--Macaulay, unless the option \texttt{AssumeCM} is set to \texttt{true}. If the ring
is not Cohen--Macaulay then \texttt{false} is returned. Next, the algorithm
computes the test module $M \subseteq \omega_{R}$ and checks to see if
$\omega_{R} \subseteq M$;  see the next section for the description of a test module.

@@

In \cite{KatzmanParameterTestIdealOfCMRings} it is shown that as $R/I$ is Cohen--Macaulay, $u$
can be taken as the generator of the cyclic module $(I^{[p]}:I) \cap (\Omega^{[p]}:\Omega)/I^{[p]}$
where $\Omega \subseteq R$ is an ideal whose image in $R/I$ is a canonical module of that ring.
